---
title: "R Notebook"
output: html_notebook
---
 
```{r Functions}
#________S_____________________
S <- function(series,k, lag){
  
  if( lag <0 || lag >= length(series)){
    stop("Invalid Lag")
  }
  
  sum = 0
  
  #Sum over all products at this lag
  for (i in 1:(length(series)-lag)){
    sum = sum + (series[i]*series[i+lag])
  }
  
  return(sum)
}

#______Huber-Phi Function________
phi <- function(X,k){
  output <- numeric(length(X))
  for (i in 1:length(X)){
    if (X[i] < -k){
      output[i] <- -k
    }
    else if (abs(X[i]) <= k){
      output[i] <- X[i]
    }
    else{
      output[i] <- k
    }
  }
  return(output)
}

#_____Transform_________________
Y_hat <- function(X,k){
  m <- mean(X)
  sigma <- sd(X)
  
  Y_hat <- (X-m)/sigma
  Y_hat <- phi(Y_hat,k)
}

#____________kernel____________
kernel <- function(x) {
  if (abs(x) <= 0.5) {
    return(1)
  } 
  else if (0.5 < abs(x) && abs(x) <= 1) {
    return(2 - 2 * abs(x))
  } 
  else {
    return(0)
  }
}
#_________sigma_hat________________
sigma_ij <- function(i,j,series,T,p,b){
  
  T_tilde <- (T-p)
  summation <- 0
  
  for (t in 1:T_tilde){
    
    t_temp <- (series[t]*series[t+j] - (1/T_tilde)*S(series,j,T_tilde))
    
    for (s in 1:T_tilde){
      
      s_temp <- (series[s]*series[s+i] - (1/T_tilde)*S(series,i,T_tilde))
      
      summation = summation + (t_temp)*(s_temp)*(kernel((s-t)/b))
      
    }
  }
  return((1/T)*summation)
}


weight_matrix <- function(letter, p, series, T, b) {
  # Takes 'e' for 1s in the diagonal
  # 'd' for descending values in the diagonal
  # 's' for entries from kernel-based long-run variance estimator (this one may take the longest)
  
  if(letter == "e") {
    W <- diag(p + 1)
  } else if(letter == "d") {
    w <- numeric(p + 1)
    for(i in 1:(p + 1)) {
      w[i] <- 1 - (i / p)
    }
    W <- diag(w)
  } else if(letter == "s") {
    w <- numeric(p + 1)
    for(i in 1:(p + 1)) {
      w[i] <- sigma_ij(i, i, series, T, p, b) 
    }
    W <- diag(w)
  } else {
    stop("Invalid letter input. Choose 'e', 'd', or 's'.")
  }
  
  return(W)
}


#_____________Test_Statistic______________
R_tilde <- function(V, W, T_tilde, T) {
  # Initialize the result list to store test statistics
  R_list <- numeric(T_tilde)
  
  # Begin calculating the test statistic
  for (k in 1:T_tilde) {
    # Calculate the first part of the test statistic
    V_sum_k <- colSums(V[1:k, , drop = FALSE])
    V_sum_T_tilde <- colSums(V[1:T_tilde, , drop = FALSE])
    
    V_hat <- V_sum_k - (k / T_tilde) * V_sum_T_tilde
    
    # Ensure V_hat is a row vector (1-row matrix) for matrix multiplication
    V_hat <- matrix(V_hat, nrow = 1)
    
    # Calculate the test statistic
    test_t <- V_hat %*% W %*% t(V_hat)
    
    # Store the test statistic
    R_list[k] <- test_t
  }
  
  # Calculate the final test statistic
  R_T <- (1 / T) * max(R_list)
  
  return(R_T)
}

#_________________Simulation__________________
simulation <- function(T,p,iterations){
  # Step 1: Generate independent standard normal variables
  T_tilde <- T - p

  # Initialize the list to store test statistics
  test_statistics <- numeric(iterations)

  # Main loop
  for (i in 1:iterations) {
    rannum <- rnorm((p + 1) * T_tilde)
    Z <- matrix(rannum, nrow = T_tilde, ncol = p + 1)
  
    # Step 2: Generate sigma and use it to find V
    b <- (T_tilde)^(1/3)
    sigma_hat <- matrix(NA, nrow = p + 1, ncol = p + 1)
  
  for (i in 1:(p + 1)) {
    for (j in 1:(p + 1)) {
      sigma_hat[i, j] <- sigma_ij(i, j, rannum, T, p, b)
    }
  }
  
  # Find the nearest positive definite matrix
  nearest_pd_matrix <- nearPD(sigma_hat)
  pd_matrix <- nearest_pd_matrix$mat
  L <- chol(pd_matrix)
  
  V <- Z %*% L
  V <- as.array(V)
  
  # Step 3: Calculate the weighted test statistic
  
  W <- diag(p + 1)
  result <- R_tilde(V, W, T_tilde, T)
  test_statistics[iteration] <- result
  
  }
  return(test_statistics)
}


#______________Run_Simulations_____________________
run_simulation <- function(T, p, num_iterations,weight_l, weight_params) {
  #weight_params should be a list containing series, T, and b in that order
  
  T_tilde <- T - p
  
  # Initialize the list to store test statistics
  test_statistics <- numeric(num_iterations)
  
  # Initialize the progress bar
  pb <- progress_bar$new(
    format = "  Progress [:bar] :percent",
    total = num_iterations,
    clear = FALSE,
    width = 60
  )
  
  # Main loop
  for (iteration in 1:num_iterations) {
    rannum <- rnorm((p + 1) * T_tilde)
    Z <- matrix(rannum, nrow = T_tilde, ncol = p + 1)
    
    # Step 2: Generate sigma and use it to find V
    b <- (T - p)^(1/3)
    sigma_hat <- matrix(NA, nrow = p + 1, ncol = p + 1)
    
    for (i in 1:(p + 1)) {
      for (j in 1:(p + 1)) {
        sigma_hat[i, j] <- sigma_ij(i, j, rannum, T, p, b)
      }
    }
    
    # Find the nearest positive definite matrix
    nearest_pd_matrix <- nearPD(sigma_hat)
    pd_matrix <- nearest_pd_matrix$mat
    L <- chol(pd_matrix)
    
    V <- Z %*% L
    V <- as.array(V)
    
    #calculate R, do this differently if given weight params
    if (is.null(weight_params)){
      W <- weight_matrix(weight_l, p)
    }
    elseif(!is.null(weight_params)){
      W <- weight_matrix(weight_l,p,weight_params[1],weight_params[2],weight_params[3])      
    }
    
    
    # Step 3: Calculate the weighted test statistic
    result <- R_tilde(V, W, T_tilde, T)
    test_statistics[iteration] <- result
    
    # Update the progress bar
    pb$tick()
  }
  
  return(test_statistics)
}
```
 
```{r}
library(Matrix)
library(stats)
library(progress)
library(openxlsx)
```

```{r Defining T, p and Z}
T <- 128
p <- 3

rho = 0.8

#define our time series here
ran_values <- rnorm(128)
AR1_series<- numeric((T-p)*(p+1))

AR1_series[1:128] <- ran_values

for (t in 129:((T-p)*(p+1))){
  AR1_series[t] <- rho*AR1_series[t-1] +rnorm(1)  
}

#We should normalize our function here
Y_hat <- Y_hat(AR1_series, k=1.5)

#Now we want to force these into our matrix Z
Z <- matrix(Y_hat, nrow=(T-p), ncol=p+1)



```

```{r}
#Now we try to generate our matrix sigma 
b = (T-p)**(1/3)
sigma_hat <- matrix(NA,nrow=(p+1),ncol=(p+1))

for (i in 1:(p+1)){
  for (j in 1:(p+1)){
    sigma_hat[i,j] <- sigma_ij(i,j,Y_hat,T,p,b)
  }
}
```

```{r}
# Find the nearest positive definite matrix
nearest_pd_matrix <- nearPD(sigma_hat)

# Extract the positive definite matrix
pd_matrix <- nearest_pd_matrix$mat

L = chol(pd_matrix);
V = Z %*% L;
V = as.array(V);
```

```{r}
W = diag(p+1)

result <- R_tilde(V, W, T_tilde, T)
print(result)
```

```{r Algorithm}
#We do the above process but now we loop over different values of rho and T

for (rho in c(0,0.8)){
  
  for (T in c(128,256,512)){
    
    #Step 1: Define and time series and make matrix Z
    
    p <- 3

    rho = 0.8

    #define our time series here
    ran_values <- rnorm(128)
    AR1_series<- numeric((T-p)*(p+1))

    AR1_series[1:128] <- ran_values

    for (t in 129:((T-p)*(p+1))){
      AR1_series[t] <- rho*AR1_series[t-1] +rnorm(1)  
    }

    #We should normalize our function here
    Y_hat <- Y_hat(AR1_series, k=1.5)

    #Now we want to force these into our matrix Z
    Z <- matrix(Y_hat, nrow=(T-p), ncol=p+1)

    #Step two: Find Matrix Sigma_hat
    b = (T-p)**(1/3)
    sigma_hat <- matrix(NA,nrow=(p+1),ncol=(p+1))

    for (i in 1:(p+1)){
      for (j in 1:(p+1)){
        sigma_hat[i,j] <- sigma_ij(i,j,Y_hat,T,p,b)
      }
    }
    
    #Step three: Find the Cholesky Decomposition L and define V
    
    # Find the nearest positive definite matrix
    nearest_pd_matrix <- nearPD(sigma_hat)

    #Extract the positive definite matrix
    pd_matrix <- nearest_pd_matrix$mat

    L = chol(pd_matrix);
    V = Z %*% L;
    V = as.array(V);
    
    #Final Step: Compute test statistic and display results
    W = diag(p+1)
    result <- R_tilde(V, W, T_tilde, T)
    print(paste("At T=",T,"and rho=",rho,"R_hat gives",result))
  }
  
}

```

```{r}
#________S_____________________
S <- function(series, k, lag){
  if (lag < 0 || lag >= length(series)){
    stop("Invalid Lag")
  }
  
  sum = 0
  # Sum over all products at this lag
  for (i in 1:(length(series) - lag)){
    sum = sum + (series[i] * series[i + lag])
  }
  return(sum)
}

#______Huber-Phi Function________
phi <- function(X, k){
  output <- numeric(length(X))
  for (i in 1:length(X)){
    if (X[i] < -k){
      output[i] <- -k
    } else if (abs(X[i]) <= k){
      output[i] <- X[i]
    } else {
      output[i] <- k
    }
  }
  return(output)
}

#_____Transform_________________
Y_hat_function <- function(X, k){
  m <- mean(X)
  sigma <- sd(X)
  
  Y_hat <- (X - m) / sigma
  Y_hat <- phi(Y_hat, k)
  return(Y_hat)
}

#____________kernel____________
kernel <- function(x) {
  if (abs(x) <= 0.5) {
    return(1)
  } else if (0.5 < abs(x) && abs(x) <= 1) {
    return(2 - 2 * abs(x))
  } else {
    return(0)
  }
}

#_________sigma_hat________________
sigma_ij <- function(i, j, series, T, p, b){
  T_hat <- (T - p)
  summation <- 0
  
  for (t in 1:T_hat){
    t_temp <- (series[t] * series[t + j] - (1 / T_hat) * S(series, j, T_hat))
    for (s in 1:T_hat){
      s_temp <- (series[s] * series[s + i] - (1 / T_hat) * S(series, i, T_hat))
      summation = summation + (t_temp) * (s_temp) * (kernel((s - t) / b))
    }
  }
  return((1 / T) * summation)
}

#_____________Test_Statistic______________
R_tilde <- function(V, W, T_tilde, T) {
  R_list <- numeric(T_tilde)
  for (k in 1:T_tilde) {
    V_sum_k <- colSums(V[1:k, , drop = FALSE])
    V_sum_T_tilde <- colSums(V[1:T_tilde, , drop = FALSE])
    V_hat <- V_sum_k - (k / T_tilde) * V_sum_T_tilde
    V_hat <- matrix(V_hat, nrow = 1)
    test_t <- V_hat %*% W %*% t(V_hat)
    R_list[k] <- test_t
  }
  R_T <- (1 / T) * max(R_list)
  return(R_T)
}

# Loop over different values of rho and T
for (rho in c(0, 0.8)){
  for (T in c(128, 256, 512)){
    # Step 1: Define and time series and make matrix Z
    p <- 3
    # rho is already defined in the loop condition, no need to redefine it
    # define our time series here
    ran_values <- rnorm(128)
    AR1_series <- numeric((T - p) * (p + 1))
    AR1_series[1:128] <- ran_values
    for (t in 129:((T - p) * (p + 1))){
      AR1_series[t] <- rho * AR1_series[t - 1] + rnorm(1)  
    }
    
    # Normalize our function here
    Y_hat_values <- Y_hat_function(AR1_series, k = 1.5)
    
    # Now we want to force these into our matrix Z
    Z <- matrix(Y_hat_values, nrow = (T - p), ncol = p + 1)
    
    # Step two: Find Matrix Sigma_hat
    b <- (T - p)^(1/3)
    sigma_hat <- matrix(NA, nrow = (p + 1), ncol = (p + 1))
    for (i in 1:(p + 1)){
      for (j in 1:(p + 1)){
        sigma_hat[i, j] <- sigma_ij(i, j, Y_hat_values, T, p, b)
      }
    }
    
    # Step three: Find the Cholesky Decomposition L and define V
    nearest_pd_matrix <- nearPD(sigma_hat)
    pd_matrix <- nearest_pd_matrix$mat
    L <- chol(pd_matrix)
    V <- Z %*% L
    V <- as.array(V)
    
    # Final Step: Compute test statistic and display results
    W <- diag(p + 1)
    T_tilde <- T - p
    result <- R_tilde(V, W, T_tilde, T)
    print(paste("At T =", T, "and rho =", rho, "R_hat gives", result))
  }
}

```

```{r}
#Step 1: Generate independent standard normal variabels
T = 128
p = 3

T_tilde = T-p


#We do this 1000 times and store the values in a list for later
test_statistics <- numeric(1000)

for (i in 1:1000){
  rannum= rnorm((p+1)*(T_tilde))

  Z <- matrix(rannum, nrow=(T_tilde), ncol=(p+1))
  
  #Step 2: Generate sigma and use to find V
  b = (T-p)**(1/3)
  sigma_hat <- matrix(NA,nrow=(p+1),ncol=(p+1))
  for (i in 1:(p+1)){
    for (j in 1:(p+1)){
      sigma_hat[i,j] <- sigma_ij(i,j,rannum,T,p,b)
    }
  }
  # Find the nearest positive definite matrix
  nearest_pd_matrix <- nearPD(sigma_hat)

  # Extract the positive definite matrix
  pd_matrix <- nearest_pd_matrix$mat

  L = chol(pd_matrix);
  V = Z %*% L;
  V = as.array(V);
  #Step 3: Calculate the weighted test statistic
  W = diag(p+1)

  result <- R_tilde(V, W, T_tilde, T)
  test_statistics[i] = result
}
```

```{r}


# Define your functions: S, phi, Y_hat, kernel, sigma_ij, and R_tilde
# Make sure these are correctly defined and available in your environment

# Step 1: Generate independent standard normal variables
T <- 128
p <- 3
T_tilde <- T - p

# Initialize the list to store test statistics
test_statistics <- numeric(1000)

# Initialize the progress bar
pb <- progress_bar$new(
  format = "  Progress [:bar] :percent in :elapsed",
  total = 1000,
  clear = FALSE,
  width = 60
)

# Main loop
for (iteration in 1:1000) {
  rannum <- rnorm((p + 1) * T_tilde)
  Z <- matrix(rannum, nrow = T_tilde, ncol = p + 1)
  
  # Step 2: Generate sigma and use it to find V
  b <- (T - p)^(1/3)
  sigma_hat <- matrix(NA, nrow = p + 1, ncol = p + 1)
  
  for (i in 1:(p + 1)) {
    for (j in 1:(p + 1)) {
      sigma_hat[i, j] <- sigma_ij(i, j, rannum, T, p, b)
    }
  }
  
  # Find the nearest positive definite matrix
  nearest_pd_matrix <- nearPD(sigma_hat)
  pd_matrix <- nearest_pd_matrix$mat
  L <- chol(pd_matrix)
  
  V <- Z %*% L
  V <- as.array(V)
  
  # Step 3: Calculate the weighted test statistic
  W <- diag(p + 1)
  result <- R_tilde(V, W, T_tilde, T)
  test_statistics[iteration] <- result
  
  # Update the progress bar
  pb$tick()
}

# Print the resulting test statistics
print(test_statistics)

```

```{r}
output <- #make an empty with 9 columns and 1000 rows
counter = 1
  
for(T in c(128,256,512)){
  for(p in c(1,3,5)){
    
    # Initialize the progress bar
  pb <- progress_bar$new(
  format = "  Progress [:bar] :percent in :elapsed",
  total = 1000,
  clear = FALSE,
  width = 60
  )
  
  temp_results <- simulation(T,p,1000)
  
  #add these results to column no. counter
  
  #Change the name of this column to (T=,p=)
  
  counter <- counter +1 
  
  pb$tick()
    
  }
}

#Finally export this dataframe as a spreadsheet

```

```{r}
# Define your functions: S, phi, Y_hat, kernel, sigma_ij, and R_tilde
# Make sure these are correctly defined and available in your environment

# Initialize an empty dataframe with 1000 rows and 9 columns
output <- data.frame(matrix(nrow = 1000, ncol = 9))

counter <- 1

for (T in c(128, 256, 512)) {
  for (p in c(1, 3, 5)) {
    
    # Initialize the progress bar
    pb <- progress_bar$new(
      format = "  Progress [:bar] :percent",
      total = 1000,
      clear = FALSE,
      width = 60
    )
    
    # Run the simulation
    temp_results <- run_simulation(T, p, 1000)
    
    # Add these results to the column number counter
    output[, counter] <- temp_results
    
    # Change the name of this column to (T=,p=)
    colnames(output)[counter] <- paste0("T=", T, ", p=", p)
    
    counter <- counter + 1
    
    pb$tick()
    
  }
}

# Export this dataframe as a spreadsheet
write.xlsx(output, "simulation_results.xlsx")

```

```{r}
# Export this dataframe as a spreadsheet
write.xlsx(output, "C://Users/filli/OneDrive/Documents/R/simulation_results.xlsx")
```

```{r}
# Define your functions: S, phi, Y_hat, kernel, sigma_ij, and R_tilde
# Make sure these are correctly defined and available in your environment

# Open our existing dataframe
output <- read.xlsx("C://Users/filli/OneDrive/Documents/R/simulation_results.xlsx")


counter <- 8
T=512

for (p in c(3, 5)) {
    
  # Run the simulation
  temp_results <- run_simulation(T, p, 1000)
  
  # Add these results to the column number counter
  output[, counter] <- temp_results
  
  # Change the name of this column to (T=,p=)
  colnames(output)[counter] <- paste0("T=", T, ", p=", p)
  
  counter <- counter + 1
  
  pb$tick()
  
}


```

```{r}
```{r}
# Define your functions: S, phi, Y_hat, kernel, sigma_ij, and R_tilde
# Make sure these are correctly defined and available in your environment

# Initialize an empty dataframe with 1000 rows and 9 columns
output <- data.frame(matrix(nrow = 1000, ncol = 27))

for (T in c(100,300,500)){
  
  for(letter in c("e","d","s")){
    
    for(p in c(1,3,5){
      
      #initialize progress bar
      # Initialize the progress bar
    pb <- progress_bar$new(
      format = "  Progress [:bar] :percent",
      total = 1000,
      clear = FALSE,
      width = 60
    )
      #run this simulation
    if (letter != "s"){
      temp_results <- run_simulation(T, p, 1000,letter)
    }
    elif(letter = "s"){
      temp_results <- run_simulations(T,p,1000,letter,weight_params=c())
    }

      #add results to output df
      
    pb$tick()

    }
    
  }
  
}

# Export this dataframe as a spreadsheet
write.xlsx(output, "simulation_results2.xlsx")

```
```